Benchmarking figures

### 2016-04-28 scripts/figures.Rmd
jupped the gun on figures again
sra info

  mutate(roc.df, exp.freq=find_freq.v(samp))-> roc.df
}

@@

  roc.df.adj<-ddply(roc.df,~samp,adjust.coords,sum.df,possible_tp,possible_vars)
  roc.df.adj<-roc.df.adj[order(roc.df.adj$adj.sensitivity),]
  mutate(roc.df.adj, gc = as.numeric(sub(pat,"\\1",samp)),
       exp.freq = sub(pat,"\\2",samp)
  )-> roc.df.adj
  fills<-ddply(roc.df.adj,~samp,fill_in_plots) # take out these lines to not fill in the plots
  roc.df.adj<-rbind(fills,roc.df.adj) # here too
  roc.df.adj$exp.freq<-as.character(roc.df.adj$exp.freq) # so that teh calculations below work since it is a factor here
  mutate(roc.df.adj,exp.freq=ifelse(test = grepl("0",exp.freq),yes=as.numeric(exp.freq)/1000,no = as.numeric(exp.freq)/100))->roc.df.adj
  roc.df.adj$exp.freq<-as.factor(roc.df.adj$exp.freq)
  roc.df.adj$exp.freq <- factor(roc.df.adj$exp.freq, levels = rev(levels(roc.df.adj$exp.freq)))
  return(roc.df.adj)
}

@@

sum.df<-read.csv("../data/process/2015-6-23/Variants/all.sum.csv",stringsAsFactors = F,comment.char = "#") # reading in the variants from the pipeline output
mfe.df<-read.csv("../data/reference/mfe.csv") 
# Getting genome lenght and segment information from the wsn33 fasta file
reference.fasta<-"../data/reference/wsn33_wt_plasmid.fa" 
segments <- fasta.seqlengths(reference.fasta)
regions.bed <- data.frame(chr = gsub("[ ].*","", names(segments)), start=12, stop=segments-13, row.names=NULL) # the univeral primers are 12 and 13 bp long
regions.bed<-mutate(regions.bed,chr=as.character(chr))

@@

sum.df<-subset(sum.df,exp.freq>0.001) # remove this to include exp.freq 0.01%
# Id the variants as T or F 
true_snv<-read.csv("../data/reference/mutant_id.csv",stringsAsFactor=F) # get the T variants
sum.df<-mutate(sum.df,category=mutation %in% true_snv$mutant) # add Column for True and false variants
wt1_mut<-subset(sum.df,Id=="WT1" & freq.var>0.01,select=mutation)
wt2_mut<-subset(sum.df,Id=="WT2"& freq.var>0.01 ,select=mutation)

@@

@@

@@

@@

Figure 5. The accuracy of the Lofreq algorithm on our 10^5% genomes/$\mu$l populations. A)  The accuracy of Lofreq using standard parameters. The specificity of the caller was scaled to account for the same number of tests as the DeepSNV algorithm. B) A table summarizing the accuracy of the ROC A using the standard cut offs applied by the algorithm.

Figure 6. The accuracy of the frequency measurements for the known variants in the 10 $^5$ genomes/$\mu$l samples. The median of each distribution is shown as a black bar.

@@

euc<-function(x,ref){
  x$Id="test"
  ref$Id="ref"
  all<-rbind(subset(x,select=c(mutation,freq.var,Id)),subset(ref,select=c(mutation,freq.var,Id)))
  tab<-as.matrix(subset(dcast(all,mutation ~ Id, value.var = "freq.var"),select=-c(mutation)))
  tab[is.na(tab)]<-0
  dist(t(tab),"manhattan")[1]
}
get_data<-function(x,setting){
  if(length(unique(x$freq.var))==1){
    exp.freq=unique(x$freq.var) # this is the perfect condition where there is only 1 frequency and it is all the same
  }else{
  exp.freq=unique(x$exp.freq)
  }
  out<-data.frame(Richness=round(rich(x),digits = 0), Entropy=format(H(x),scientific = T,digits = 3),'L1-norm'=round(euc(x,subset(perfect,freq.var==exp.freq)),digits=3)) #"1/D"=D(x),
  rownames(out)<-setting
  return(t(out))
}
require(xtable)
tabulate<-function(input,freq){
x<-data.frame()
for(i in 1:length(freq)){  
get_data(subset(perfect,freq.var==freq[i]),"Expected")->perf
get_data(subset(sum.df,gc==input & exp.freq==freq[i] & p.val<0.01 & category %in% c(T,F)),"DeepSNV")->deepSNV
get_data(subset(lofreq.df,gc==input & exp.freq==freq[i] & category %in% c(T,F)),"Lofreq")->lo
get_data(subset(cut.read.df.all,gc==input & exp.freq==freq[i] & p.val<0.01 & category %in% c(T,F) ),"DeepSNV modified")->deepSNV.f
y<-cbind(perf,deepSNV.f,lo,deepSNV)
x<-rbind(x,y)
}
names<-data.frame(Frequency=c("","5.0%","","","1.0%","","","0.5%",""),names=rep(c("Richness","Entropy","L1-norm"),times=3))
x<-cbind(names,x)
#x<-subset(x,select=c(names,Expected,DeepSNV,Lofreq,DeepSNV Final))
x<-rename(x,c("names"=""))
xtab<-xtable(x)
align(xtab) <- rep("c", 7)
print(xtab,include.rownames = FALSE,hline.after=c(-1,0,3,6))
}

@@

```{r,results="asis",cache=F}
tabulate(5,c(0.05,0.01,0.005))

@@

sum.df<-subset(sum.df,exp.freq>0.001) # remove this to include exp.freq 0.01%
# Id the variants as T or F 
true_snv<-read.csv("../data/reference/mutant_id.csv",stringsAsFactor=F) # get the T variants
sum.df<-mutate(sum.df,category=mutation %in% true_snv$mutant) # add Column for True and false variants
wt1_mut<-subset(sum.df,Id=="WT1" & freq.var>0.01,select=mutation)
wt2_mut<-subset(sum.df,Id=="WT2"& freq.var>0.01 ,select=mutation)

@@

>>>>

  mutate(roc.df, exp.freq=find_freq.v(samp))-> roc.df # this yeilds null in the hiseq.roc function below
}

@@

  roc.df.adj<-ddply(roc.df,~samp,adjust.coords,sum.df,possible_tp,possible_vars)
  roc.df.adj<-roc.df.adj[order(roc.df.adj$adj.sensitivity),]
  mutate(roc.df.adj, gc = as.numeric(sub(pat,"\\1",samp)), # This fills in the null column given in the roc_df function above. line 161
       exp.freq = sub(pat,"\\2",samp)
  )-> roc.df.adj
  fills<-ddply(roc.df.adj,~samp,fill_in_plots) # take out these lines to not fill in the plots
  roc.df.adj<-rbind(fills,roc.df.adj) # here too
  roc.df.adj$exp.freq<-as.character(roc.df.adj$exp.freq) # so that teh calculations below work since it is a factor here
  mutate(roc.df.adj,exp.freq=ifelse(test = grepl("0",exp.freq),yes=as.numeric(exp.freq)/1000,no = as.numeric(exp.freq)/100))->roc.df.adj
  roc.df.adj$exp.freq<-as.factor(roc.df.adj$exp.freq)
  roc.df.adj$exp.freq <- factor(roc.df.adj$exp.freq, levels = rev(levels(roc.df.adj$exp.freq)))
  return(roc.df.adj)
}

@@

sum.df<-read.csv("../data/process/2015-6-23/Variants/all.sum.csv",stringsAsFactors = F,comment.char = "#") # reading in the variants from the pipeline output
mfe.df<-read.csv("../data/reference/mfe.csv") 
# Getting genome lenght and segment information from the wsn33 fasta file
reference.fasta<-"../data/reference/wsn33_wt_plasmid.fa" 
segments <- fasta.seqlengths(reference.fasta)
regions.bed <- data.frame(chr = gsub("[ ].*","", names(segments)), start=12, stop=segments-13, row.names=NULL) # the univeral primers are 12 and 13 bp long
regions.bed<-mutate(regions.bed,chr=as.character(chr))
# The samples are named in the following format log(genome copy/ul)_expected frequency of variants We'll use that information here to get the gc# and exp.freq for each variant called

@@

sum.df<-subset(sum.df,exp.freq>0.001) # remove this to include exp.freq 0.01%
# Id the variants as T or F 
sum.df<-ddply(sum.df,~chr,primer.cut) # removes any variants outside of priming sites (in this case it removes 0)

true_snv<-read.csv("../data/reference/mutant_id.csv",stringsAsFactor=F) # get the T variants
sum.df<-mutate(sum.df,category=mutation %in% true_snv$mutant) # add Column for True and false variants
wt1_mut<-subset(sum.df,Id=="WT1" & freq.var>0.01,select=mutation)
wt2_mut<-subset(sum.df,Id=="WT2" &freq.var>0.01 ,select=mutation)

wt_mut<-intersect(wt1_mut$mutation,wt2_mut$mutation) # there aren't any of these

@@

possible_vars<-total_possible_vars-length(wt_mut)-dim(mfe.df)[1]-20 # removing the mfe and wt_mut sites
hiseq_vars<-possible_vars
print(hiseq_vars)
hiseq_kb<-sum(regions.bed$length)/1000

@@

mapq<-ggplot(subset(sig.df,category %in% c(T,F) & gc==5),aes(x=MapQ,fill=category))+geom_histogram(position="dodge")+theme(legend.position="none")+scale_fill_manual(name="",labels=c("False positives","True positives"),values=c("#b1bdc5ff","#08519cff"))+ylab("SNV count")

phred<-ggplot(subset(sig.df,category %in% c(T,F) & gc==5),aes(x=Phred,fill=category))+geom_histogram(position="dodge")+theme(legend.position="none")+scale_fill_manual(name="",labels=c("False positives","True positives"),values=c("#b1bdc5ff","#08519cff"))+scale_x_continuous(limits = c(30,40),breaks = c(30,32.5,35,37.5,40))+ylab("SNV count")

pdf("../results/figures/mapq.pdf")
mapq
dev.off()

pdf("../results/figures/phred.pdf")
phred
dev.off()

@@

Figure 5. The accuracy of the frequency measurements for the known variants in the 10 $^5$ genomes/$\mu$l samples. The median of each distribution is shown as a black bar.

@@

Figure 6. The accuracy of the Lofreq algorithm on our 10^5^ genomes/$\mu$l populations. A)  The accuracy of Lofreq using standard parameters. The specificity of the caller was scaled to account for the same number of tests as the DeepSNV algorithm. B) A table summarizing the accuracy of the ROC A using the standard cut offs applied by the algorithm.

@@

euc<-function(x,ref){
  x$Id="test"
  ref$Id="ref"
  all<-rbind(subset(x,select=c(mutation,freq.var,Id)),subset(ref,select=c(mutation,freq.var,Id)))
  tab<-as.matrix(subset(dcast(all,mutation ~ Id, value.var = "freq.var"),select=-c(mutation)))
  tab[is.na(tab)]<-0
  dist(t(tab),"manhattan")[1]
}
get_data<-function(x,setting){
  if(length(unique(x$freq.var))==1){
    exp.freq=unique(x$freq.var) # this is the perfect condition where there is only 1 frequency and it is all the same
  }else{
  exp.freq=unique(x$exp.freq)
  }
  out<-data.frame(Richness=round(rich(x),digits = 0), Entropy=format(H(x),scientific = T,digits = 3),'L1-norm'=round(euc(x,subset(perfect,freq.var==exp.freq)),digits=3)) #"1/D"=D(x),
  rownames(out)<-setting
  return(t(out))
}
require(xtable)
tabulate<-function(input,freq){
x<-data.frame()
for(i in 1:length(freq)){  
get_data(subset(perfect,freq.var==freq[i]),"Expected")->perf
get_data(subset(sum.df,gc==input & exp.freq==freq[i] & p.val<0.01 & category %in% c(T,F)),"DeepSNV")->deepSNV
get_data(subset(lofreq.df,gc==input & exp.freq==freq[i] & category %in% c(T,F)),"Lofreq")->lo
get_data(subset(cut.read.df.all,gc==input & exp.freq==freq[i] & p.val<0.01 & category %in% c(T,F) ),"DeepSNV modified")->deepSNV.f
y<-cbind(perf,deepSNV.f,lo,deepSNV)
x<-rbind(x,y)
}
names<-data.frame(Frequency=c("","5.0%","","","1.0%","","","0.5%",""),names=rep(c("Richness","Entropy","L1-norm"),times=3))
x<-cbind(names,x)
#x<-subset(x,select=c(names,Expected,DeepSNV,Lofreq,DeepSNV Final))
x<-rename(x,c("names"=""))
write.csv(x,file="../results/tables/table_1.csv")
xtab<-xtable(x)
align(xtab) <- rep("c", 7)
print(xtab,include.rownames = FALSE,hline.after=c(-1,0,3,6))

}

@@

```{r,results="asis",cache=F}
tabulate(5,c(0.05,0.01,0.005))
#

@@

sum.df<-subset(sum.df,exp.freq>0.001) # remove this to include exp.freq 0.01%
# Id the variants as T or F 
true_snv<-read.csv("../data/reference/mutant_id.csv",stringsAsFactor=F) # get the T variants
sum.df<-mutate(sum.df,category=mutation %in% true_snv$mutant) # add Column for True and false variants
sum.df<-ddply(sum.df,~chr,primer.cut)

wt1_mut<-subset(sum.df,Id=="WT1" & freq.var>0.01,select=mutation)
wt2_mut<-subset(sum.df,Id=="WT2"& freq.var>0.01 ,select=mutation)

@@

```{r,table2}
data<-read.csv("../data/reference/patient.csv")
data<-subset(data,p.val<0.01 & freq.var<0.9)
data<-ddply(data,~chr,primer.cut) # removes sites out side of primers

data.cut<-subset(data,Phred>35 & MapQ>30 & Read_pos<94 & Read_pos>32)

possible_vars<-total_possible_vars
get_data<-function(x){
  out<-data.frame(Richness=round(rich(x),digits = 0), Entropy=format(H(x),scientific = T,digits = 3)) #"1/D"=D(x),
  #rownames(out)<-setting
  return(out)
}
get_data.cut<-function(x){
  out<-data.frame(Richness.mod=round(rich(x),digits = 0), Entropy.mod=format(H(x),scientific = T,digits = 3)) #"1/D"=D(x),
  #rownames(out)<-setting
  return(out)
}

x<-ddply(data,~Id+DPI+FINAL_RESULT+season+log10(gc_ul),get_data)

y<-ddply(data.cut,~Id+DPI+log10(gc_ul),get_data.cut)

x<-join(x,y)
x<-subset(x,!(Id%in% c(1240,1241)))
x<-x[order(x$`log10(gc_ul)`,decreasing = T),]
write.csv(x,"../results/tables/table_2.cali_vic.csv")
ggplot(data,aes(pos,freq.var))+geom_point()+facet_wrap(~chr)

ggplot(data.cut,aes(pos,freq.var))+geom_point()+facet_wrap(~chr)

ggplot(data,aes(freq.var))+geom_histogram()+scale_x_log10()

ggplot(data.cut,aes(freq.var))+geom_histogram()+scale_x_log10()+scale_y_log10()

x<-subset(data.cut,Id %in% c(1376,1401,1405,1374,1227,1321,1229,1245))
```
